---
title: "p8105_mtp_yc4018"
author: "Yuxuan Chen"
date: "10/28/2021"
output: github_document
---
## Introduction 


## Problem 1 – Data.

```{r message = FALSE}
library(tidyverse)
library(patchwork)

#set the theme of all graphs
theme_set(theme_minimal() + theme(legend.position = "right"))

#set color of all graphs
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

1. load, clean, and tidy the mtp_df data
```{r}
# Based on the article, the ages were divided into age groups of 18–30, 31–40, 41–50, 51–60, and >61.

mtp_df = readxl::read_excel("./data/p8105_mtp_data.xlsx", range = "A9:I1230") %>%  #read the table without header info
  janitor::clean_names() %>% 
  mutate(
    sex = recode(sex, `1` = "male", `0` = "female"),
    eop_size_mm = if_else(is.na(eop_size_mm), 0, eop_size_mm), # replace NA with 0 in `eop_size_mm` variable. 
    age = as.integer(age),
    age_group = case_when(
      age_group %in% c("6","7","8") ~ "6+",
      TRUE   ~ age_group)) %>% 
  rename(eop_size_group = eop_size, 
         fhp_group = fhp_category) %>% 
  mutate(
      age_group = fct_relevel(age_group, "1", "2", "3", "4", "5", "6+"),
      eop_size_group = fct_relevel(eop_size_group, "0", "1", "2", "3", "4", "5"),
      eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2")),
      fhp_group = fct_relevel(fhp_group,"0", "1", "2", "3", "4", "5","6", "7")
  )
knitr::kable(mtp_df[0:10,])
```

 - After loading the dataset, I cleaned up the variable names and converted them to lower snake cases. Based on header information, I changed the observations of `1` and `0` to `male` and `female` respectively, and  replaced NA with 0 in `eop_size_mm` variable. Next I changed the observations in `age` variable to be integer and combined the ages older than 60 as the group of `6+`. I renamed the `eop_size` to `eop_size_group`, `fhp_category` to `fhp_group` for naming consistency. Lastly, I ordered the factors for `age_group`,`eop_size_group`,`eop_visibility_classification`, and `fhp_group` variables.
 
\newline
 - The size of the resulting dataset is  (`r dim(mtp_df)`), which contains 1221 observations of 9 variables. The key variables are `age_group`, `eop_size_group`, `eop_visibility_classification`, and `fhp_group`, because all these variables categorized data into different groups based on different criteria. 

```{r message = FALSE}
n_ptcp = mtp_df %>% count()

age_sex_df = 
  mtp_df %>% 
  group_by(age_group, sex) %>%
  mutate(
    age_group = recode(
      age_group, "1" = "<18", "2" = "18–30", "3" = "31–40", "4" = "41–50", "5" = "51–60", "6+" = ">60")) %>% 
  summarize(n_obs = n())

age_sex_df %>% 
  pivot_wider(
    names_from = sex,
    values_from = n_obs) %>% 
  knitr::kable(caption = "Age and Gender Distribution")

age_sex_df %>% 
  ggplot(aes(x = age_group, y = n_obs, group = sex, color = sex)) + 
  geom_point(alpha = .5) +
  geom_line(alpha = .5) +
  labs(
    title = "Scatter Plot of Age Distribution Based on Sex",
    x = "Age group",
    y = "Number of Observations"
  )

```

  - There are `r n_ptcp` participants included. Male and female have the similar distributions of number of observations in each age group.

2. Issues in the data: 
```{r}
y = as.integer(c(18, 31, 41, 51, 61))
class(mtp_df$age)
# age:
mtp_df %>% 
  ggplot(aes(x = age_group, y = age, color = age_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(18, 31, 41, 51, 61), linetype = 3) +
  labs(
    title = "Scatter Plot of Age Data",
    x = "Age Group (years)",
    y = "Age") +
  scale_x_discrete(
    labels = c("<18", "18–30", "31–40", "41–50", "51–60", ">60")) +
  scale_color_hue(
    labels = c("<18", "18–30", "31–40", "41–50", "51–60", ">60")) +
  annotate("text", x = 0.7, y = 18, label = "y=18", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 30, label = "y=31", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 41, label = "y=41", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 51, label = "y=51", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 61, label = "y=61", vjust = -1, size = 3)

mtp_df %>% 
  filter(age_group == 1) %>% 
  select(age, age_group) %>% 
  knitr::kable()
```

 - There are 2 data placed in the <18 age group, especially one with 45 age is mistakenly placed in this age group. 

```{r}
#EOP sizes group
mtp_df %>% 
  ggplot(aes(x = eop_size_group, y = eop_size_mm, color = eop_size_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(5, 10, 15, 20, 25), linetype = 3) +
  labs(
    title = "Scatter Plot of EOP Sizes in EOP Sizes Groups",
    x = "EOP Size Group",
    y = "EOP Size (mm)") +
  scale_color_hue(
    name = "EOP size group",
    labels = c("0 = 0-5mm", "1 = 5-10mm", "2 = 10-15mm", "3 = 15-20mm", "4 = 20-25mm", "5 = 25+mm"))

mtp_df %>% 
  filter(!eop_size_group %in% c("0","1","2","3","4","5")) %>% 
  select(eop_size_group, eop_size_mm) %>% 
  knitr::kable()
```
 
 - There is one data with 15mm EOP size was mistakenly placed in 14.6 size group. Also there are many misclassified data in size groups of 1, 3, 4, 5.

```{r}
#eop_visibility_classification
mtp_df %>% 
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = eop_visibility_classification)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(5), linetype = 3) +
  labs(
    title = "Scatter Plot of EOP Sizes in EOP Visibility Classification Groups",
    x = "EOP Visibility Classification Group",
    y = "EOP Size (mm)") +
  scale_color_hue(
    name = "EOP size group",
    labels = c("0 = 0mm", "1 = 0-5mm", "2 => 5mm"))
```

 - There are a lot misclassified data.

```{r}
mtp_df %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = fhp_group, y = fhp_size_mm, color = fhp_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(10, 20, 30, 40, 50, 60, 70, 80), linetype = 3) +
  labs(
    title = "Scatter Plot of FHP Sizes",
    x = "FHP Size Group",
    y = "FHP Size (mm)") +
  scale_color_hue(
    name = "FHP size group",
    labels = c("0 = 0-10mm", "1 = 10-20mm", "2 = 20-30mm","3 = 30-40mm", "4 = 40-50mm", "5 = 50-60mm", "6 = 60-70mm", "7 = 70-80mm"))

mtp_df %>% 
  filter(!fhp_group %in% c("0","1","2","3","4","5", "6","7")) %>% 
  select(fhp_group,fhp_size_mm) %>% 
  knitr::kable()
```
 - There is one data with 30.3mm FHP size was mistakenly placed in 30.8 size group. Also there are many misclassified data in size groups 1, 2, and 4.

## Problem 2 – Visualization.

```{r message = FALSE}
FHP_plot = 
  mtp_df %>% 
  filter(age_group != 1) %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot(alpha = .5) +
  labs(
    title = "FHP Values Across the Age Groups and Sexes",
    x = "Age Group (years)",
    y = "FHP Size (mm)") +
  scale_x_discrete(labels = c("18-30 years", "30s", "40s", "50s", ">60")) + 
  stat_summary(fun = mean, geom = "point", size = 1, position = position_dodge(width = 0.75)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.3, position = position_dodge(width = 0.75))

eeop_plot = 
  mtp_df %>% 
  mutate(
    eeop = case_when(
      eop_size_group %in% c("0", "1") ~ "no",
      eop_size_group %in% c("2", "3", "4", "5") ~ "yes",
      TRUE   ~ "")) %>% 
  group_by(age_group, sex, eeop) %>% 
  summarize(n_obs = n()) %>% 
  mutate(rate = prop.table(n_obs)) %>% 
  filter(eeop == "yes") %>% 
  ggplot(aes(x = age_group, y = rate)) + 
  geom_point() +  
  geom_line(aes(group = sex, color = sex))  +
  labs(
    title = "Distribution of the Rate of EEOP In Each Age Group",
    x = "Age Group (years)",
    y = "EEOP prop") +
  scale_x_discrete(labels = c("18-30 years", "30s", "40s", "50s", ">60"))
FHP_plot / eeop_plot
```

 - Improved fig.3 shows the distribution of FHP data across the age groups and sexes. In the 60+ age group, both male and female have the highest FHP size. Male has higher FHP sizes than female within all age groups since the median of FHP sizes of male are higher. Both mean and sd are close to median in each age and sex group.

 \newline.  
 - Bottom plot shows the distribution of rate of EEOP in each age and sex group. The trends of distribution in improved fig.4 are similar to the original plot. Male has higher EEOP rate than female. 18-30 years group has the highest EEOP rate. With age increases, EEOP rate decreases and EEOP becomes unlikely to occur. After the age of 40s, EEOP rate slightly increases 
 
```{r message = FALSE}
mtp_df %>%
  filter(age_group != 1) %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = sex)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "black", size = 0.5, se = FALSE) +
  facet_grid(
    sex ~ age_group,
    labeller = labeller(age_group = c("2" = "18-30 years", "3" = "30s", "4" = "40s", "5" = "50s", "6+" = ">60s"))) +
  labs(
    title = "Distribution between FHP Size and EOP Size",
    caption = "Association between FHP size and EOP size in each age and sex group",
    x = "FHP Size (mm)",
    y = "EOP Size (mm)")
```

In both sex groups, EOP size and FHP size have positive relationship in every age group. 

## Problem 3 – Reproducing reported results.

```{r}
mtp_df %>% 
  group_by(age_group) %>% 
  summarize(n = n()) %>% 
  mutate(
    age_group = recode(
      age_group, "1" = "<18", "2" = "18–30", "3" = "31–40", "4" = "41–50", "5" = "51–60", "6+" = ">60")) %>% 
  knitr::kable(col.names = c("Age Group", "Sample Size"))
```

1. No. The authors’ stated sample size was: 18–30: n = 300, 31–40: n = 200, 41–50: n = 200, 51–60: n = 200, and >60: n = 300, which is different with the data I have.

```{r}
mean(pull(mtp_df, fhp_size_mm), na.rm = TRUE) 
mtp_df %>% 
  group_by(sex) %>% 
  summarize(
    mean = mean(fhp_size_mm, na.rm = TRUE),
    sd = sd(fhp_size_mm, na.rm = TRUE)
  ) %>% 
  knitr::kable()
```

2. The reported overall mean FHP was 26 ±mm, for male was 28±15mm, and for female was 24±11mm, which are consistent with the data I have available.

```{r}
mtp_df %>% 
  filter(
    age_group != 1,
    eop_size_group %in% c("0", "1", "2", "3", "4", "5")) %>% 
  mutate(
    eeop = case_when(
      eop_size_group %in% c("0", "1") ~ "no",
      eop_size_group %in% c("2", "3", "4", "5") ~ "yes",
      TRUE   ~ "")) %>% 
  group_by(eeop) %>% 
  summarize(n_obs = n()) %>% 
  mutate(rate = prop.table(n_obs)) %>% 
  knitr::kable()
```

3. EEOP is when the bony outgrowth exceeds 10mm and can be evaluated by `eop_size_mm` or `eop_size_group`. The calculated prevalence of EEOP equals 32% which is smaller than authors' finding.

```{r}
mtp_df %>% 
  filter(age_group != 1) %>% 
  group_by(age_group) %>% 
  count(fhp_group) %>% 
  mutate(rate = prop.table(n)) %>% 
  filter(fhp_group %in% c('4', '5', '6', '7')) %>% 
  summarize(sum = sum(rate)) %>% 
  knitr::kable(col.names = c("Age Group", "Frequency of FHP >40 mm"))
```

4. The broad trend is consistent with my data, but my frequency of FHP>40mm in 60+ age group is 32.5%, which is smaller than the finding by author.

## Problem 4 – Discussion.
 - From my calculation, increasing in age resulted increasing in FHP size, which is positively related to EOP size. Hence, the high EEOP rate in 18-30 years group do recommends a potential avenue for prevention intervention through posture improvement education in this cohort. However, there are many mistakes in data and many misclassified data in original report, which makes the quality of the results doubtful. Some calculated values, like prevalence of EEOP, are also different with authors' findings. Even though the broad trends of the original report's conclusions are consistent with mine, we cannot conclude that cell phones are causing horn growth since poor posture may not directly cause EEOP formation, and there may be other poor postures but not just using cell phones. Hence, prevalence of top poor postures and time of cell phone usage among different age groups may be helpful to address the hypothesis.
