---
title: "p8105_mtp_yc4018"
author: "Yuxuan Chen"
date: "10/28/2021"
output: github_document
---

## Introduction 
   In this project, the author correction data was used to re-analysis the relationship of how FHP size, age, and sex affect EOP size and re-conduct the hypothesis of whether cell phones are causing horn growth. By doing this to test the consistency and quality of the results in the original report. 

```{r message = FALSE}
library(tidyverse)
library(patchwork)

#set the theme of all graphs
theme_set(theme_minimal() + theme(legend.position = "right"))

#set color of all graphs
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1 – Data.
```{r}
# load, clean, and tidy mtp_df data
## Based on the article, the ages were divided into age groups of 18–30, 31–40, 41–50, 51–60, and >61.

mtp_df = readxl::read_excel("./data/p8105_mtp_data.xlsx", range = "A9:I1230") %>%  #read the table without header info
  janitor::clean_names() %>% 
  mutate(
    sex = recode(sex, `1` = "male", `0` = "female"),
    eop_size_mm = if_else(is.na(eop_size_mm), 0, eop_size_mm), # replace NA with 0 in `eop_size_mm` variable. 
    age = as.integer(age),
    age_group = case_when(
      age_group %in% c("6","7","8") ~ "6+",
      TRUE   ~ age_group)) %>% 
  rename(eop_size_group = eop_size, 
         fhp_group = fhp_category) %>% 
  mutate(
      age_group = fct_relevel(age_group, "1", "2", "3", "4", "5", "6+"),
      eop_size_group = fct_relevel(eop_size_group, "0", "1", "2", "3", "4", "5"),
      eop_visibility_classification = factor(eop_visibility_classification, levels = c("0", "1", "2")),
      fhp_group = fct_relevel(fhp_group,"0", "1", "2", "3", "4", "5","6", "7")
  )
knitr::kable(mtp_df[0:10,])
```


```{r message = FALSE, collapse=TRUE}
n_ptcp = mtp_df %>% count()

age_sex_df = 
  mtp_df %>% 
  group_by(age_group, sex) %>%
  mutate(
    age_group = recode(
      age_group, "1" = "<18", "2" = "18–30", "3" = "31–40", "4" = "41–50", "5" = "51–60", "6+" = ">60")) %>% 
  summarize(n_obs = n())

age_sex_df %>% 
  pivot_wider(
    names_from = sex,
    values_from = n_obs) %>% 
  knitr::kable(caption = "Age and Gender Distribution")

age_sex_df %>% 
  ggplot(aes(x = age_group, y = n_obs, group = sex, color = sex)) + 
  geom_point(alpha = .5) +
  geom_line(alpha = .5) +
  labs(
    title = "Scatter Plot of Age Distribution Based on Sex",
    x = "Age group",
    y = "Number of Observations"
  )

```

1. After loading the dataset, I cleaned up the variable names and converted them to lower snake cases. Based on header information, I changed the observations of `1` and `0` to `male` and `female` respectively and replaced NA with 0 in `eop_size_mm`. Next, I changed the observations in `age` to integer and combined the ages older than 60 as the group of `6+`. I renamed the `eop_size` to `eop_size_group`, `fhp_category` to `fhp_group` for naming consistency. Lastly, I ordered the factors of `age_group`,`eop_size_group`,`eop_visibility_classification`, and `fhp_group` variables.    
 - The size of the resulting dataset is  (`r dim(mtp_df)`), which contains 1221 observations of 9 variables. The key variables are `age_group`, `eop_size_group`, `eop_visibility_classification`, and `fhp_group`, because all these variables categorized data into different groups based on different criteria. 
 - There are `r n_ptcp` participants included. Male and female have similar distributions of the number of observations in each age group.

2. Issues in the data: 
```{r}
# age:
mtp_df %>% 
  ggplot(aes(x = age_group, y = age, color = age_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(18, 31, 41, 51, 61), linetype = 3) +
  labs(
    title = "Scatter Plot of Age Data",
    x = "Age Group (years)",
    y = "Age") +
  scale_color_hue(
    name = "Age group",
    labels = c("1 = <18 years", "2 = 18–30 years", "3 = 31–40 years", "4 = 41–50 years", "5 = 51–60 years", "6+ = >60 years")) +
  annotate("text", x = 0.7, y = 18, label = "y=18", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 30, label = "y=31", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 41, label = "y=41", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 51, label = "y=51", vjust = -1, size = 3) +
  annotate("text", x = 0.7, y = 61, label = "y=61", vjust = -1, size = 3)

mtp_df %>% 
  filter(age_group == 1) %>% 
  select(age, age_group) %>% 
  knitr::kable()
```

 - There are 2 data placed in the <18 age group. One data with 45 age is misclassified into this age group. 

```{r}
#EOP sizes group
mtp_df %>% 
  ggplot(aes(x = eop_size_group, y = eop_size_mm, color = eop_size_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(5, 10, 15, 20, 25), linetype = 3) +
  labs(
    title = "Scatter Plot of EOP Sizes in EOP Size Groups",
    x = "EOP Size Group",
    y = "EOP Size (mm)") +
  scale_color_hue(
    name = "EOP size group",
    labels = c("0 = 0-5mm", "1 = 5-10mm", "2 = 10-15mm", "3 = 15-20mm", "4 = 20-25mm", "5 = 25+mm"))

mtp_df %>% 
  filter(!eop_size_group %in% c("0","1","2","3","4","5")) %>% 
  select(eop_size_mm, eop_size_group) %>% 
  knitr::kable()
```
 
 - There is one data with 15mm EOP size is misclassified into in 14.6 size group. Also, there are many misclassified data in 1, 3, 4, 5 size groups.

```{r}
#eop_visibility_classification
mtp_df %>% 
  ggplot(aes(x = eop_visibility_classification, y = eop_size_mm, color = eop_visibility_classification)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(5), linetype = 3) +
  labs(
    title = "Scatter Plot of EOP Sizes in EOP Visibility Classification Groups",
    x = "EOP Visibility Classification Group",
    y = "EOP Size (mm)") +
  scale_color_hue(
    name = "EOP size group",
    labels = c("0 = 0mm", "1 = 0-5mm", "2 => 5mm"))
```

 - There are a lot misclassified data.

```{r}
mtp_df %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = fhp_group, y = fhp_size_mm, color = fhp_group)) +
  geom_point(alpha = .5) +
  geom_hline(yintercept = c(10, 20, 30, 40, 50, 60, 70, 80), linetype = 3) +
  labs(
    title = "Scatter Plot of FHP Sizes",
    x = "FHP Size Group",
    y = "FHP Size (mm)") +
  scale_color_hue(
    name = "FHP size group",
    labels = c("0 = 0-10mm", "1 = 10-20mm", "2 = 20-30mm","3 = 30-40mm", "4 = 40-50mm", "5 = 50-60mm", "6 = 60-70mm", "7 = 70-80mm"))

mtp_df %>% 
  filter(!fhp_group %in% c("0","1","2","3","4","5", "6","7")) %>% 
  select(fhp_group,fhp_size_mm) %>% 
  knitr::kable()
```

 - There is one data with 30.3mm FHP size is misclassified into 30.8 size group. Also, there are many misclassified data in size groups 1, 2, and 4.

## Problem 2 – Visualization.

```{r message = FALSE}
FHP_plot = 
  mtp_df %>% 
  filter(age_group != 1) %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = age_group, y = fhp_size_mm, fill = sex)) +
  geom_boxplot(alpha = .5) +
  labs(
    title = "FHP Values Across the Age Groups and Sexes",
    x = "Age Group (years)",
    y = "FHP Size (mm)") +
  scale_x_discrete(labels = c("18-30 years", "30s", "40s", "50s", ">60")) + 
  stat_summary(fun = mean, geom = "point", size = 1, position = position_dodge(width = 0.75)) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.3, position = position_dodge(width = 0.75))

eeop_plot = 
  mtp_df %>% 
  mutate(
    eeop = case_when(
      eop_size_group %in% c("0", "1") ~ "no",
      eop_size_group %in% c("2", "3", "4", "5") ~ "yes",
      TRUE   ~ "")) %>% 
  group_by(age_group, sex, eeop) %>% 
  summarize(n_obs = n()) %>% 
  mutate(rate = prop.table(n_obs)) %>% 
  filter(eeop == "yes") %>% 
  ggplot(aes(x = age_group, y = rate)) + 
  geom_point() +  
  geom_line(aes(group = sex, color = sex))  +
  labs(
    title = "Distribution of the Rate of EEOP In Each Age Group",
    x = "Age Group (years)",
    y = "EEOP prop") +
  scale_x_discrete(labels = c("18-30 years", "30s", "40s", "50s", ">60"))
FHP_plot / eeop_plot
```

 - Improved fig.3 shows the distribution of FHP data across the age groups and sexes. With increasing age, FHP size also increases. In the 60+ age group, both male and female have the highest FHP sizes. Male has higher FHP sizes than female within all age groups since the median of FHP sizes of the male is higher. Both mean and sd are close to the median in each age and sex group.    
 - Bottom plot shows the distribution of the rate of EEOP in each age and sex group. The trends of distribution in improved fig.4 are similar to the original plot. Male has a higher EEOP rate than female. The 18-30 years group has the highest EEOP rate. With age increases, the EEOP rate decreases, and EEOP becomes unlikely to occur. After the age of 40s, the EEOP rate increases again.
 
```{r message = FALSE}
mtp_df %>%
  filter(age_group != 1) %>% 
  drop_na(fhp_size_mm) %>% 
  ggplot(aes(x = fhp_size_mm, y = eop_size_mm, color = sex)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "black", size = 0.5, se = FALSE) +
  facet_grid(
    sex ~ age_group,
    labeller = labeller(age_group = c("2" = "18-30 years", "3" = "30s", "4" = "40s", "5" = "50s", "6+" = ">60s"))) +
  labs(
    title = "Distribution between FHP Size and EOP Size",
    caption = "Association between FHP size and EOP size in each age and sex group",
    x = "FHP Size (mm)",
    y = "EOP Size (mm)")
```

 - In both sex groups, FHP size and EOP size have a positive relationship in every age group. In each age and sex group, with increasing of FHP Size will result increasing of EOP Size.

## Problem 3 – Reproducing reported results.

```{r}
mtp_df %>% 
  group_by(age_group) %>% 
  summarize(n = n()) %>% 
  mutate(
    age_group = recode(
      age_group, "1" = "<18", "2" = "18–30", "3" = "31–40", "4" = "41–50", "5" = "51–60", "6+" = ">60")) %>% 
  knitr::kable(col.names = c("Age Group", "Sample Size"))
```

1. No, since the authors’ stated sample sizes are: 18–30: n = 300, 31–40: n = 200, 41–50: n = 200, 51–60: n = 200, and >60: n = 300, which are different with the data in the calculated table above.

```{r}
# overall mean FHP
mean(pull(mtp_df, fhp_size_mm), na.rm = TRUE)

mtp_df %>% 
  group_by(sex) %>% 
  summarize(
    mean = mean(fhp_size_mm, na.rm = TRUE),
    sd = sd(fhp_size_mm, na.rm = TRUE)
  ) %>% 
  knitr::kable()
```

2. Yes, since the authors stated overall mean FHP was 26 ±mm, the male's mean FHP was 28±15mm and the female's mean FHP was 24±11mm, which is consistent with the data in the calculated table above.

```{r}
# Because of misclassification, I used `eop_size_mm` to distinguish the size of 10mm. 
mtp_df %>% 
  filter(
    age_group != 1,
    eop_size_group %in% c("0", "1", "2", "3", "4", "5")) %>% 
  mutate(
    eeop = case_when(
      eop_size_mm <= 10 ~ "no",
      eop_size_mm > 10 ~ "yes",
      TRUE   ~ "")) %>% 
  group_by(eeop) %>% 
  summarize(n_obs = n()) %>% 
  mutate(rate = prop.table(n_obs)) %>% 
  knitr::kable()
```

3. EEOP is named when the size of bony outgrowth exceeds 10mm, which can be evaluated by `eop_size_mm` and `eop_size_group`. The calculated prevalence of EEOP equals 32% which is smaller than the authors’ finding.

```{r}
trend_df = 
  mtp_df %>% 
  filter(age_group != 1) %>% 
  group_by(age_group) %>% 
  count(fhp_group) %>% 
  mutate(rate = prop.table(n)) %>% 
  filter(fhp_group %in% c('4', '5', '6', '7')) %>% 
  summarize(sum = sum(rate))

knitr::kable(trend_df, col.names = c("Age Group", "Frequency of FHP >40 mm"))
trend_df %>% 
  ggplot(aes(x = age_group, y = sum, group = NA)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Distribution of Frequency of FHP > 40mm",
    x = "Age Group (years",
    y = "Frequency of FHP > 40mm")

```

4. The broad trend is consistent with my data, but my frequency of FHP>40mm in the 60+ age group is 32.5%, which is smaller than the finding by the authors.

## Problem 4 – Discussion.

 - In summary, increasing in age will result increasing in FHP size which is positively related to EOP size. The high EEOP rate in the 18-30 years group does recommend a potential avenue for prevention intervention through posture improvement education. However, there are many mistakes in data and many misclassified data in the original report, which makes the quality of the results doubtful. Some calculated values are also different from the authors' findings. Even the broad trends from the original report are consistent with mine, we cannot conclude that cell phones causes horn growth since poor postures may not directly cause EEOP formation, and there may exist other poor postures. Hence, the prevalence of top poor postures and time of cell phone usage among different age groups may be helpful to address the hypothesis.
